Data Management for LSST Special Programs

### 2023-11-08 sec_intro.tex
update section titles
define Main Survey, add labels

\section{Executive Summary} \label{sec:execsum}

@@

\item All visits should have a region label \textit{and} a program 
label, and these labels should be propagated to processed images and catalog data.
(Section~\ref{ssec:sci_labels})

@@

\section{Introduction}\label{sec:intro}

@@

\subsection{Wide Fast Deep (WFD) Program}

The WFD is the core science program of the LSST, designed to achieve the science 
goals defined by the Science Requirements Document (SRD; \citeds{LPM-17}).

The WFD program is defined from a scheduling perspective as a certain observing strategy
(i.e., a set of cadences and filter balances) applied to a low-dust region 
(extragalactic areas) and several special regions:
the Galactic bulge and plane; the South Celestial Pole (SCP); the North Ecliptic Spur (NES);
and the Virgo Cluster \citedsp{PSTN-055}.

The WFD is expected to use alternative standard visits in the $u$-band and 
standard visits in all other bands \citedsp{PSTN-055}.
Different cadences, filter balances, rotational dithers, etc., might
be applied in the different regions composing WFD.
For example, rolling cadence is expected to be implemented in the low-dust region only.

WFD program data will be processed with regular (non-special) processing.
This processing will produce, among other data products (LSE-163), 
a contiguous sky footprint that covers at 
least $\sim$18000 deg$^2$ with $\gtrsim$800 visits per field, 
and is expected to be accomplished with 85--90\% of the observing time.
This contiguous all-sky "deep coadd" may have region-dependent processing parameters and inputs
(e.g., different calibration parameters or deblending algorithms for crowded fields)
and will be of variable depth, as most special regions recieve fewer visits.

@@

This is a term used only in this document to refer to the image processing 
described in the Data Products Definitions Document
(LSE-163) that are designed for, and will be applied to, the WFD program's observations.

@@

This is a Rubin Data Management term used to refer to sky regions within
or beyond the WFD footprint which have a distinct observing strategy from
the WFD program's plans for that region (e.g., different visit type, cadence, filter distribution).

Special Programs are typically driven by specific science goals that build on or 
add to the core science pillars of the LSST.
They include non-WFD LSST components such as the Deep Drilling 
Fields (DDFs) and the mini-, micro-, and nano-surveys \citedsp{PSTN-055}.
About 10--15\% of the total 10-year LSST will be spent obtaining 
observations associated with Special Programs.

Data products from Special Programs is subject to the Rubin Data Policy \citedsp{RDO-013}
in the same way as data products from the WFD program.

@@

This is a Rubin Data Management term to describe processing that 
uses components of the LSST Science Pipelines and is applied by Rubin
Data Management to images from Special Programs.
Special Processing creates data products that are unique and separate from those produced
by regular processing for the WFD program.

@@

The sky areas of mini-surveys can be within, adjacent to, or detached from the WFD footprint.
Mini-surveys can have non-standard visits.
Target-of-opportunity (TOO) observations for, e.g., the discovery of optical counterparts to 
multi-messenger astrophysical phenomena, are considered a type of mini-survey in this document.
For a list of the mini-surveys under consideration,
see \citeds{PSTN-055}.

>>>>

\section{Memoranda} \label{sec:mem}

@@

\item All visits should have a region label \textit{and} an observing mode
label, and these labels should be propagated to processed images and catalog data.
(Section~\ref{ssec:sci_labels})

@@

\section{Terms and Definitions}\label{sec:intro}

@@

\subsection{Region Label}

Based on Right Acension and Declination.

To be applied to data by Rubin Data Management, e.g., using a look-up table.

Region labels would include, e.g., WFD low-dust, Galactic Plane, Galactic 
Bulge, SCP, NES, or Virgo; DDF fieldname; mini-survey region name.

\subsection{Observing Mode Label}

Based on the scheduler mode at the time of the observation
(i.e., survey cadence or strategy).

To be applied by the Survey Scheduler team via the observations scheduler.

Observing mode labels might include, e.g., WFD, DDF, TOO, mini-survey name, or other 
options like engineering, commissioning, and director's discretionary.

\subsection{Main Survey and Wide-Fast-Deep (WFD)}

The LSST Main Survey includes the low-dust Wide-Fast-Deep (WFD) area 
(which includes low-dust extragalactic and Galactic regions) 
and several special regions:
the dusty areas of the Galactic bulge and plane; the South Celestial Pole (SCP); 
the North Ecliptic Spur (NES); and the Virgo Cluster \citedsp{PSTN-055}.
The WFD component is the core component of the LSST Main Survey, designed to achieve the science 
goals defined by the Science Requirements Document (SRD; \citeds{LPM-17}).

The Main Survey will be executed with a set of observing modes, and 
is expected to use alternative standard visits in the $u$-band and 
standard visits in all other bands \citedsp{PSTN-055}.
Different cadences, filter balances, rotational dithers, etc., 
might be applied in the different regions composing Main Survey.
For example, rolling cadence is expected to be implemented in the low-dust 
WFD region only.

Data from the Main Survey will be processed with regular (non-special) processing.
This processing will produce, among other data products (LSE-163), 
a contiguous sky footprint that covers at 
least $\sim$18000 deg$^2$ (with $\gtrsim$800 visits per field in WFD regions), 
and is expected to be accomplished with 85--90\% of the observing time.
This contiguous all-sky "deep coadd" may have region-dependent processing parameters and inputs
(e.g., different calibration parameters or deblending algorithms for crowded fields)
and will be of variable depth, as most special regions recieve fewer visits.

@@

This is a term used only in this document to refer to the image processing 
described in the Data Products Definitions Document
(LSE-163) that are designed for, and will be applied to, the Main Survey's observations.

@@

This is a Rubin Data Management term used to refer to sky regions within
or beyond the Main Survey footprint which have a distinct observing strategy from
the Main Survey's plans for that region (e.g., different visit type, cadence, filter distribution).

Special Programs are typically driven by specific science goals that build on or 
add to the core science pillars of the LSST.
They include LSST components such as the Deep Drilling 
Fields (DDFs) and the mini-, micro-, and nano-surveys \citedsp{PSTN-055}.
About 10--15\% of the total 10-year LSST will be spent obtaining 
observations associated with Special Programs.

Data products from Special Programs is subject to the Rubin Data Policy \citedsp{RDO-013}
in the same way as data products from the Main Survey.

@@

This is a Rubin Data Management term to describe processing that 
uses components of the LSST Science Pipelines and is applied by Rubin
Data Management to images from Special Programs.
Special Processing creates data products that are unique and separate from those produced
by regular processing for the Main Survey.

@@

The sky areas of mini-surveys can be within, adjacent to, or detached from the Main Survey footprint.
Mini-surveys can have non-standard visits.
Target-of-opportunity (TOO) observations for, e.g., the discovery of optical counterparts to 
multi-messenger astrophysical phenomena, are considered a type of mini-survey in this document.
For a list of the mini-surveys under consideration,
see \citeds{PSTN-055}.
